R Notebook

### 2020-02-17 analysis/final_analysis.Rmd
make figs

## Model-based analyses

@@

```{r}
d.onDistractor <- exp2_d %>% 
  left_join(read_csv('../data/final/experiment2/mouseOverDistractor/distractorMouseFromMongo.csv') %>% 
              mutate(critical = ifelse(critical == 'True', 1, 0))
            ) %>%
  select(-serverTime, -X1) %>%
  filter(attemptNum == 0) %>%
  filter(correct == 1) %>%
  filter(critical == 1) %>%
  group_by(gameid, trialNum) %>%
  mutate(toggle = floor((row_number() - 1) / 2)) %>%
  mutate(onOff = ifelse(is.na(onOff), 'off', onOff)) %>%
  spread(onOff, timeElapsed) %>%
  mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
  group_by(gameid,condition,objectSet, trialType) %>%
  summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
  mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor))) %>%
  mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
  ungroup()

dodge <- position_dodge(width=0.9)
d.onDistractor %>%
  group_by(condition, trialType) %>%
  tidyboot_mean(column = totalAmtTimeOnDistractor) %>%
  mutate(trial = ifelse(trialType == 'base', 'baseline', 'experimental')) %>%
  ggplot(aes(x = condition, y=empirical_stat, fill = condition, group = trial)) +
    geom_bar(aes(alpha = trial), stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
    theme_few() +
    scale_alpha_manual(values=c(.25, .75)) +
    ylab('(log) ms hovering over distractor') +
    scale_fill_colorblind()+
    guides(fill='none') +
    theme(aspect.ratio = 2, legend.position = 'top')

@@

```{r}
summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 +trialType|gameid) + (1 +trialType|objectSet), 
             data = d.onDistractor))
```

@@

```{r}
summary(glmer(hovered ~ trialType * condition + (1|gameid) + (1+trialType|objectSet), data = d.onDistractor, family = 'binomial'))
```
## Speaker analyses

@@

>>>>

## Model-based analyses (Appendix)

### Resource-rational simulations

```{r}
RR <- bind_rows(
  read_csv('../modeling/output/RRlistenerOutput.csv') %>% mutate(src = 'listener'),
  read_csv('../modeling/output/RRspeakerOutput.csv') %>% mutate(src ='speaker')
)

ggplot(RR, aes(x = weighting, y = prob, color = perspectiveCost, group = perspectiveCost)) +
  geom_line() +
  facet_wrap(~ src) +
  theme_few() +
  xlim(0,1) +
  theme(aspect.ratio = 1)

ggsave('../../pragmatics_and_perspective_taking_tex/cognitive_science_revision/figures/simulations.pdf')
```

### Dynamics simulations

```{r}
dynamics <- bind_rows(
  read_csv('../modeling/output/dynamicsOutput.csv')
)

dynamics %>%
  group_by(numDatapoints) %>%
  filter(prob == max(prob)) %>%
  ggplot(aes(x = numDatapoints, y = weighting)) +
    geom_line() +
    theme_few() +
    xlim(0, 30) +
    ylim(.5, 1) +
    labs(x = 'length of interaction', y = 'optimal weighting') +
    theme(aspect.ratio = 1)

ggsave('../../pragmatics_and_perspective_taking_tex/cognitive_science_revision/figures/dynamics.pdf', width = 2, height = 2,
       units = 'in')
```

### Model comparison

@@

```{r}
d.onDistractor.correctOnly <- exp2_d %>% 
  filter(correct) %>%
  filter(attemptNum == 0) %>%
  filter(critical == 1) %>%
  left_join(read_csv('../data/final/experiment2/mouseOverDistractor/distractorMouseFromMongo.csv'),
            by = c('gameid', 'trialNum', 'objectSet', 'trialType', 'condition', 'attemptNum')) %>%
  select(-serverTime, -X1, -ends_with('.y')) %>%
  group_by(gameid, trialNum) %>%
  mutate(toggle = floor((row_number() - 1) / 2)) %>%
  mutate(onOff = ifelse(is.na(onOff), 'off', onOff)) %>%
  spread(onOff, timeElapsed) %>%
  mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
  group_by(gameid,trialNum,condition,objectSet, trialType) %>%
  summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
  mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor))) %>%
  mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
  ungroup()

dodge <- position_dodge(width=0.9)
d.onDistractor.correctOnly %>%
  group_by(condition, trialType) %>%
  tidyboot_mean(column = totalAmtTimeOnDistractor) %>%
  mutate(trial = ifelse(trialType == 'base', 'baseline', 'experimental')) %>%
  ggplot(aes(x = condition, y=empirical_stat, fill = condition, group = trial)) +
    geom_bar(aes(alpha = trial), stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
    theme_few() +
    scale_alpha_manual(values=c(.25, .75)) +
    ylab('(log) ms hovering over distractor') +
    scale_fill_colorblind()+
    guides(fill='none') +
    theme(aspect.ratio = 2, legend.position = 'top')

@@

```{r}
summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 +trialType|gameid) + (1 +trialType|objectSet), 
             data = d.onDistractor.correctOnly))
```

@@

```{r}
summary(glmer(hovered ~ trialType * condition + (1|gameid) + (1+trialType|objectSet), 
              data = d.onDistractor.correctOnly, family = 'binomial'))
```

## Adaptation over time

```{r}
critTrials %>% 
  group_by(gameid) %>% 
  mutate(trialNum = row_number()) %>% 
  group_by(trialNum) %>% 
  summarize(prop = sum(correct)/length(correct)) %>%
  #mutate(error = ifelse(correct == 0, 'error', 'no error')) %>%
  ggplot(aes(x= trialNum, y = prop)) +
    geom_point() +
    geom_smooth(method = 'lm')
```

Mixed-effects model on raw error probability

```{r}
summary(glmer(correct ~ trialNum * condition + (1 | gameid), family = 'binomial',
             data = critTrials %>%   
                    group_by(gameid) %>%
                    mutate(trialNum = row_number())))
```

Supplementary Bayesian version with more complete random effects

```{r}
library(brms)
summary(brm(correct ~ trialNum * condition  + (1 + trialNum | gameid), family = 'bernoulli', 
              data = critTrials %>%   
                    group_by(gameid) %>%
                    mutate(trialNum = row_number(),
                           lastRoundIncorrect = 1-lag(correct)), control = list(adapt_delta = .95)))
```

## Speaker analyses

@@

# Supplemental experiment

```{r}
library(tidyverse)
d.supp.clicks <- read_tsv('../data/final/supplemental_exp/clickedObj/allClicks.csv')
paste0(c('we have', length(unique(d.supp.clicks$gameid)), 'participants.'), collapse = ' ')

# Following same criterion as above, we remove participants who make more than one
# non-critical error (ignoring errors in the initial practice block)
exclusions <- d.supp.clicks %>%
  rowwise() %>%
  filter(occlusions != 'critical'  || (occlusions == 'critical' && !clickedHiddenObj && !correct)) %>%
  filter(trialNum > 4) %>%
  group_by(gameid) %>%
  summarize(nonCriticalErrors = sum(!correct)) %>%
  filter(nonCriticalErrors > 1) %>%
  pull(gameid)
```

```{r}
d.supp.clicks %>%
  filter(!(gameid %in% exclusions)) %>%
  group_by(gameid) %>%
  filter(occlusions == 'critical') %>%
  mutate(instance = row_number()) %>%
  filter(clickedHiddenObj) %>%
  group_by(gameid) %>%
  summarize(errorInstances = paste0(instance, collapse =  ',')) #%>%
  # group_by(instance) %>%
  # summarize(clickedHidden =sum(clickedHiddenObj))
```

```{r}
critical_only <- read_tsv('../data/final/supplemental_exp/clickedObj/allClicks.csv') %>%
  filter(occlusions =='critical') %>%
  rowwise() %>%
  mutate(shape_of_target = strsplit(intendedName, '_')[[1]][3]) %>%
  rowwise() %>%
  mutate(i = grep(shape_of_target, strsplit(names, ',')[[1]])[1]) %>%
  mutate(criticalDistractorX = as.numeric(gsub('\\[', '', strsplit(gridXs, ',')[[1]][i])),
         criticalDistractorY = as.numeric(gsub('\\[', '', strsplit(gridYs, ',')[[1]][i])),
         criticalDistractorLeftBorder = 600 * (criticalDistractorX - 1),
         criticalDistractorRightBorder = 600 * (criticalDistractorX),
         criticalDistractorTopBorder = 600 * (criticalDistractorY-1),
         criticalDistractorBottomBorder = 600 * (criticalDistractorY)) %>%
  select(gameid, trialNum, criticalDistractorLeftBorder,criticalDistractorRightBorder,criticalDistractorTopBorder,criticalDistractorBottomBorder)
```

```{r}
distractor <- left_join(
  read_tsv('../data/final/supplemental_exp/updateMouse/mouse.csv'),
   read_tsv('../data/final/supplemental_exp/clickedObj/allClicks.csv'),
  by = c('trialNum', 'gameid')) %>%
  filter(occlusions == 'critical') %>%
  left_join(critical_only, by = c('gameid', 'trialNum')) %>%
  rowwise() %>%
  mutate(overDistractor = (
    rawMouseX < criticalDistractorRightBorder && 
    rawMouseX > criticalDistractorLeftBorder &&
    rawMouseY < criticalDistractorBottomBorder && 
    rawMouseY > criticalDistractorTopBorder
    )
  ) %>%
  filter(!(gameid %in% exclusions)) %>%
  group_by(gameid, trialNum) %>% 
  summarize(hovered = sum(overDistractor) > 0) %>%
  mutate(instance = row_number())

distractor %>%
  group_by(instance) %>%
  tidyboot_mean(sum(hovered) / length(hovered)) %>%
  ggplot(aes(x = instance, y = empirical_stat)) +
    geom_point() +
    #geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0) +
    geom_smooth(method = 'lm', formula = y ~ poly(x, 2), se = F) +
    theme_bw() +
    ylab('% trials hovering over distractor') +
    xlab('occurrence in trial sequence') +
    theme(aspect.ratio = 1/2)

ggsave('hoverReduction.pdf')
library(lme4)
library(tidyboot)
library(lmerTest)
summary(lmer(hovered ~ instance + (1 | gameid), data = distractor))
```
